---
title: "Project EDA"
author: "Douglas Byers, Matt Muller, TJ Rogers"
date: "11/2/2020"
output:
  pdf_document:
    fig_height: 2
    fig_width: 3.5
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, warning = FALSE}
library(dplyr)
library(readr)
library(mosaic)
suppressMessages(library(tidyverse))
suppressMessages(library(glmnet))
suppressMessages(library(class))
suppressMessages(library(modelr))
library(lubridate)
suppressMessages(library(caret)) ## for knn3
suppressMessages(library(vip)) ##maybe needs to be installed
games_2019 <- read_csv("~/Mscs 341b S21/Project/Matt_TJ_Doug/games_2019.csv")
games_2018 <- read_csv("~/Mscs 341b S21/Project/Matt_TJ_Doug/games_2018.csv")
```

```{r, warning = FALSE}
games_2018 <- games_2018%>%
  separate(Date, into = c("Mth", "Day"), sep = " ")%>%
  mutate(Month = if_else(Mth == "Mar", 3, 
                   if_else(Mth == "Apr", 4,
                     if_else(Mth == "May", 5,
                        if_else(Mth == "Jun", 6,
                           if_else(Mth == "Jul", 7,
                              if_else(Mth == "Aug", 8,
                                if_else(Mth == "Sep", 9, 10))))))))%>%
  unite(Date, Month, Day, Year, sep = "/")

games.train <- games_2018 %>%
  mutate(Home = if_else(is.na(Var.4), 1, 0))%>%
  separate(Rslt, into = c("Result", "Score"), sep = ",")%>%
  separate(Score, into = c("Runs Scored", "Runs Allowed"), sep = "-") %>%
  mutate(Home_Team = if_else(Home == 1, team, Opp))%>%
  mutate(Away_Team = if_else(Home == 0, team, Opp))%>%
  mutate(Matchup_1 = Home_Team)%>%
  mutate(Matchup_2 = Away_Team)%>%
  unite(Matchup, Matchup_1, Matchup_2, sep = ",")%>%
  mutate(Run_Diff = as.numeric(`Runs Scored`) - as.numeric(`Runs Allowed`))%>%
  mutate(Win = ifelse(Result == "W", 1, 0))%>%
  arrange(mdy(Date), Matchup, Home)

g1 <- games.train%>%
  filter(team == lag(team) & Opp == lag(Opp))

games.train <- setdiff(games.train, g1)

games_2019 <- games_2019%>%
  separate(Date, into = c("Mth", "Day"), sep = " ")%>%
  mutate(Month = if_else(Mth == "Mar", 3, 
                   if_else(Mth == "Apr", 4,
                     if_else(Mth == "May", 5,
                        if_else(Mth == "Jun", 6,
                           if_else(Mth == "Jul", 7,
                              if_else(Mth == "Aug", 8,
                                if_else(Mth == "Sep", 9, 10))))))))%>%
  unite(Date, Month, Day, Year, sep = "/")

games.test <- games_2019 %>%
  mutate(Home = if_else(is.na(Var.4), 1, 0))%>%
  separate(Rslt, into = c("Result", "Score"), sep = ",")%>%
  separate(Score, into = c("Runs Scored", "Runs Allowed"), sep = "-") %>%
  mutate(Home_Team = if_else(Home == 1, team, Opp))%>%
  mutate(Away_Team = if_else(Home == 0, team, Opp))%>%
  mutate(Matchup_1 = Home_Team)%>%
  mutate(Matchup_2 = Away_Team)%>%
  unite(Matchup, Matchup_1, Matchup_2, sep = ",")%>%
  mutate(Run_Diff = as.numeric(`Runs Scored`) - as.numeric(`Runs Allowed`))%>%
  mutate(Win = ifelse(Result == "W", 1, 0))%>%
  arrange(mdy(Date), Matchup, Home)


g2 <- games.test %>%
  filter(team == lag(team) & Opp == lag(Opp))

games.test <- setdiff(games.test, g2)


cols.num <- c(8:32, 36:62, 65:75, 77, 81)
games.train[cols.num] <- sapply(games.train[cols.num],as.numeric)
games.test[cols.num] <- sapply(games.test[cols.num],as.numeric)

```

Let's run some initial box and scatter plots to compare statistics we think may have an effect on a game with the outcome (win or loss).
```{r}
(RBox = ggplot (games.test, aes(y=as.factor(Win), x=R))+
  geom_boxplot())

(RABox = ggplot (games.test, aes(y=as.factor(Win), x=RA))+
  geom_boxplot())

(HBox = ggplot (games.test, aes(y=as.factor(Win), x=H))+
  geom_boxplot())

(HRBox = ggplot (games.test, aes(y=as.factor(Win), x=HR))+
  geom_boxplot())

(OBPBox = ggplot (games.test, aes(y=as.factor(Win), x=OBP))+
  geom_boxplot())

(SLGBox = ggplot (games.test, aes(y=as.factor(Win), x=SLG))+
  geom_boxplot())

(OPSBox = ggplot (games.test, aes(y=as.factor(Win), x=OPS))+
  geom_boxplot())

(HABox = ggplot (games.test, aes(y=as.factor(Win), x=HA))+
  geom_boxplot())

(BBABox = ggplot (games.test, aes(y=as.factor(Win), x=BBA))+
  geom_boxplot())

(HRABox = ggplot (games.test, aes(y=as.factor(Win), x=HRA))+
  geom_boxplot())

(StrPBox = ggplot (games.test, aes(y=as.factor(Win), x=StrPerc))+
  geom_boxplot())

(OPSABox = ggplot (games.test, aes(y=as.factor(Win), x=OPSA))+
  geom_boxplot())

(BAABox = ggplot (games.test, aes(y=as.factor(Win), x=BAA))+
  geom_boxplot())

(SOPercBox = ggplot (games.test, aes(y=as.factor(Win), x=SOPerc))+
  geom_boxplot())

(SOPercHBox = ggplot (games.test, aes(y=as.factor(Win), x=SOPercH))+
  geom_boxplot())
```

```{r}
(BA_BAAplot = ggplot(games.test)+
  geom_point(aes(x= BA, y = BAA, color = as.factor(Win)), alpha = 0.4))

(SOPplot = ggplot(games.test)+
  geom_point(aes(x= SOPercH, y = SOPerc, color = as.factor(Win)), alpha = 0.4))
```

* Quick Model Test

##Linear Model
We'll try a simple linear model with random statistics just to see how it could look.
```{r}
mod.lm.samp <- lm(Win ~ OBP + SOPerc + H + HR + WHIP + HRA, 
             data = games.train)

games.test.samp <- games.test %>%
  add_predictions(mod.lm.samp,
                  type = "response") %>%
  mutate(class = if_else(pred > 0.5, 1,0))

table(games.test.samp$Win, games.test.samp$class)
(err.lm.samp <- with(games.test.samp, mean(Win != class)))
```

```{r}
summary(mod.lm.samp)
```

## Log Model
Same process with log model
```{r}
mod.log.samp <- glm(Win ~ OBP + SOPerc + H + HR + WHIP + HRA, 
             data = games.train,
             family = "binomial")

games.test.samp.log <- games.test %>%
  add_predictions(mod.log.samp,
                  type = "response") %>%
  mutate(class = if_else(pred > 0.5, 1,0)) #class instead of response

table(games.test.samp.log$Win, games.test.samp.log$class)
(err.log.samp <- with(games.test.samp.log, mean(Win != class)))
```

```{r}
summary(mod.log.samp)
```


## Lasso Model
Now we'll work on creating a lasso model. First we want to make sure all variables included are numeric and that the data is represented as a matrix.
```{r}
games.y <- data.matrix(games.train$Win)


games.x.lasso <- games.train %>%
  select(8:32, 36:62, 65:75, 77)
games.x.lasso[is.na(games.x.lasso)] <- 0
games.x <- data.matrix(games.x.lasso)
```

```{r}
lasso.cv <- cv.glmnet(games.x,
                      games.y,
                      family="binomial",
                      type.measure="class", 
                      alpha=1)

plot(lasso.cv)
```

```{r}
vip(lasso.cv)
```

```{r}
lambda.opt <- lasso.cv$lambda.1se
id <- with(lasso.cv,which(lasso.cv$lambda==lambda.opt))
(err.lasso <- lasso.cv$cvm[id])
```


Obviously, RBI's and Runs/runs against will have an impact on the outcome. Let's look at more vague statistics to see how well they predict the model without these.

```{r}
games.test.y <- data.matrix(games.test$Win)
games.test.x1 <- games.test %>%
  select(10:11, 13:16, 18:32, 36:38, 42:62, 65:75, 77)
games.test.x1[is.na(games.test.x1)] <- 0
games.test.x <- data.matrix(games.test.x1)
```

```{r}
##New data set, gotta adjust the numbers selected

games.x.noR1 <- games.train %>%
  select(10:11, 13:16, 18:32, 36:38, 42:62, 65:75, 77)
games.x.noR1[is.na(games.x.noR1)] <- 0
games.x1 <- data.matrix(games.x.noR1)
```

```{r}
lasso.cv1 <- cv.glmnet(games.x1,
                      games.y,
                      family="binomial",
                      type.measure="class", 
                      alpha=1)

plot(lasso.cv1)
```

```{r}
vip(lasso.cv1)
```

```{r}
lambda.opt <- lasso.cv1$lambda.1se
id1 <- with(lasso.cv1,which(lasso.cv1$lambda==lambda.opt))
(err.lasso1 <- lasso.cv1$cvm[id1])
```

```{r}
preds.l <- predict(lasso.cv1, newx=games.x1,
                 type="class")
##The error rate on the testing dataset
table(games.y, preds.l)
```


Here we see that OBPA and WHIP have a lot of relative importance, with BA, BAA, Home, and OBP follow quite far behind.

## Ridge Model

We'll use the same smaller dataset that we used for lasso for ridge as well (numeric explanatory variables) to predict a ridge model, along with an optimal lambda.

```{r}
ridge.cv1 <- cv.glmnet(games.x1, games.y,
                      family="binomial",
                      type.measure="class", 
                      alpha=0)
plot(ridge.cv1)
```

```{r}
log(ridge.cv1$lambda.1se)
log(ridge.cv1$lambda.1se)
```

```{r}
lambda.opt1 <- ridge.cv1$lambda.1se
id2 <- with(ridge.cv1,which(ridge.cv1$lambda==lambda.opt1))
(err.ridge1 <- ridge.cv1$cvm[id2])
```

```{r}
preds.r <- predict(ridge.cv1, newx=games.test.x,
                 type="class")
##The error rate on the testing dataset
table(games.test.y ,preds.r)
```

```{r}
vip(ridge.cv1)
```


```{r}
mod.ridge.opt <- glmnet(games.x, games.y,
                        family="binomial",
                        type.measure="class",
                        alpha=0,
                        lambda=lambda.opt)
```

```{r}
preds <- predict(mod.ridge.opt, newx=games.test.x,
                 type="class")
##The error rate on the testing dataset
table(games.test.y ,preds)
```

```{r}
(err.ridge.opt <- mean((games.test.y != preds)))
```


Interesting, we actually see no statistic representing Runs through the ridge model which is pretty cool. The greatest importance seems to be represented by specific pitching and hitting stats that wouldn't directly correlate with wins at first sight (which runs and runs against clearly would).

These statistics include BA, OBPA, BAA, SLGA, StrPerc, SLG, SOPercH, and OPSA (where BA has clear importance over other variables, and OBPA and BAA have a significant difference in importance over the others).

* Least important features?
```{r}
res.vip <- vip(ridge.cv1,num_features=55)
vip.dat <- res.vip[["data"]]
vip.dat[52:55,]
```

CSA, Str, HBPA, and Pit have the least significance according to our ridge model.

```{r}
res.vip1 <- vip(lasso.cv1,num_features=55)
vip.dat1 <- res.vip1[["data"]]
vip.dat1[52:55,]
```

## Linear and Log model Part II

Now that we've seen these, let's try linear and log models with the reduced variables (numeric ones) that we've used.

```{r}
##New data set, gotta adjust the numbers selected
games.test1 <- games.test %>%
  select(Win, 10:11, 13:16, 18:32, 36:38, 42:62, 65:75, 77)

games.train1 <- games.train %>%
  select(Win, 10:11, 13:16, 18:32, 36:38, 42:62, 65:75, 77)

games.train1[is.na(games.train1)] <- 0
games.test1[is.na(games.test1)] <- 0
```


```{r}
mod.lm <- lm(Win ~ ., 
             data = games.train1)

games.test.lm <- games.test1 %>%
  add_predictions(mod.lm,
                  type = "response") %>%
  mutate(class = if_else(pred > 0.5, 1,0))

table(games.test.lm$Win, games.test.lm$class)
(err.lm <- with(games.test.lm, mean(Win != class)))
```

```{r}
vip(mod.lm)
```

```{r}
summary(mod.lm)
```

```{r}
res.vip2 <- vip(mod.lm,num_features=55)
vip.dat2 <- res.vip2[["data"]]
vip.dat2[52:55,]
```


```{r}
mod.log <- glm(Win ~ ., 
             data = games.train1,
             family = "binomial")

games.test.log <- games.test1 %>%
  add_predictions(mod.log,
                  type = "response") %>%
  mutate(class = if_else(pred > 0.5, 1,0))

table(games.test.log$Win, games.test.log$class)
(err.log <- with(games.test.log, mean(Win != class)))
```

```{r}
vip(mod.log)
```

```{r}
summary(mod.log)
```

```{r}
res.vip3 <- vip(mod.log,num_features=55)
vip.dat3 <- res.vip3[["data"]]
vip.dat3[52:55,]
```

The log model has an insanely low error of 0.005.


```{r}
games.train2 <- games.train%>%
  mutate(Home_Win = if_else(Home == 1 & Win == 1 | Home == 0 & Win == 0, 1, 0),
         Home_BA = if_else(Home == 0, lead(BA), BA),
         Away_BA = if_else(Home == 1, lag(BA), BA),
         Home_OBP = if_else(Home == 0, lead(OBP), OBP),
         Away_OBP = if_else(Home == 1, lag(OBP), OBP),
         Home_SLG = if_else(Home == 0, lead(SLG), SLG),
         Away_SLG = if_else(Home == 1, lag(SLG), SLG),
         Home_OPS = if_else(Home == 0, lead(OPS), OPS),
         Away_OPS = if_else(Home == 1, lag(OPS), OPS),
         Home_ISO = Home_SLG - Home_BA,
         Away_ISO = Away_SLG - Away_BA,
         Home_BAA = if_else(Home == 0, lead(BAA), BAA),
         Away_BAA = if_else(Home == 1, lag(BAA), BAA),
         Home_OBPA = if_else(Home == 0, lead(OBPA), OBPA),
         Away_OBPA = if_else(Home == 1, lag(OBPA), OBPA),
         Home_SLGA = if_else(Home == 0, lead(SLGA), SLGA),
         Away_SLGA = if_else(Home == 1, lag(SLGA), SLGA),
         Home_OPSA = if_else(Home == 0, lead(OPSA), OPSA),
         Away_OPSA = if_else(Home == 1, lag(OPSA), OPSA),
         Home_StrPerc = if_else(Home == 0, lead(StrPerc), StrPerc),
         Away_StrPerc = if_else(Home == 1, lag(StrPerc), StrPerc),
         Home_ERA = if_else(Home == 0, lead(ERA), ERA),
         Away_ERA = if_else(Home == 1, lag(ERA), ERA),
         Home_WHIP = if_else(Home == 0, lead(WHIP), WHIP),
         Away_WHIP = if_else(Home == 1, lag(WHIP), WHIP),
         Home_FIP = if_else(Home == 0, lead(FIP), FIP),
         Away_FIP = if_else(Home == 1, lag(FIP), FIP),
         Home_SOPerc = if_else(Home == 0, lead(SOPerc), SOPerc),
         Away_SOPerc = if_else(Home == 1, lag(SOPerc), SOPerc),
         Home_SOPercH = if_else(Home == 0, lead(SOPercH), SOPercH),
         Away_SOPercH = if_else(Home == 1, lag(SOPercH), SOPercH),
         Home_K9 = if_else(Home == 0, lead(K_9), K_9),
         Away_K9 = if_else(Home == 1, lag(K_9), K_9),
         Home_KBB = if_else(Home == 0, lead(K_BB), K_BB),
         Away_KBB = if_else(Home == 1, lag(K_BB), K_BB),
         Home_HR9 = if_else(Home == 0, lead(HR_9), HR_9),
         Away_HR9 = if_else(Home == 1, lag(HR_9), HR_9),
         Home_RunDiff = if_else(Home == 0, lead(Run_Diff), Run_Diff))%>%
  select(Date, Matchup, Home_Team, Home_BA, Home_OBP, Home_SLG, Home_OPS, Home_ISO, Home_SOPercH, Home_BAA, Home_OBPA, Home_SLGA, Home_OPSA, Home_StrPerc, Home_ERA, Home_WHIP, Home_FIP, Home_SOPerc, Home_KBB, Home_K9, Home_HR9, Away_Team, Away_BA, Away_OBP, Away_SLG, Away_OPS, Away_ISO, Away_SOPercH, Away_BAA, Away_OBPA, Away_SLGA, Away_OPSA, Away_StrPerc, Away_ERA, Away_WHIP, Away_FIP, Away_SOPerc, Away_KBB, Away_K9, Away_HR9, Home_Win, Home_RunDiff)%>%
  unique
```


```{r}
games.test2 <- games.test%>%
  mutate(Home_Win = if_else(Home == 1 & Win == 1 | Home == 0 & Win == 0, 1, 0),
         Home_BA = if_else(Home == 0, lead(BA), BA),
         Away_BA = if_else(Home == 1, lag(BA), BA),
         Home_OBP = if_else(Home == 0, lead(OBP), OBP),
         Away_OBP = if_else(Home == 1, lag(OBP), OBP),
         Home_SLG = if_else(Home == 0, lead(SLG), SLG),
         Away_SLG = if_else(Home == 1, lag(SLG), SLG),
         Home_OPS = if_else(Home == 0, lead(OPS), OPS),
         Away_OPS = if_else(Home == 1, lag(OPS), OPS),
         Home_ISO = Home_SLG - Home_BA,
         Away_ISO = Away_SLG - Away_BA,
         Home_BAA = if_else(Home == 0, lead(BAA), BAA),
         Away_BAA = if_else(Home == 1, lag(BAA), BAA),
         Home_OBPA = if_else(Home == 0, lead(OBPA), OBPA),
         Away_OBPA = if_else(Home == 1, lag(OBPA), OBPA),
         Home_SLGA = if_else(Home == 0, lead(SLGA), SLGA),
         Away_SLGA = if_else(Home == 1, lag(SLGA), SLGA),
         Home_OPSA = if_else(Home == 0, lead(OPSA), OPSA),
         Away_OPSA = if_else(Home == 1, lag(OPSA), OPSA),
         Home_StrPerc = if_else(Home == 0, lead(StrPerc), StrPerc),
         Away_StrPerc = if_else(Home == 1, lag(StrPerc), StrPerc),
         Home_ERA = if_else(Home == 0, lead(ERA), ERA),
         Away_ERA = if_else(Home == 1, lag(ERA), ERA),
         Home_WHIP = if_else(Home == 0, lead(WHIP), WHIP),
         Away_WHIP = if_else(Home == 1, lag(WHIP), WHIP),
         Home_FIP = if_else(Home == 0, lead(FIP), FIP),
         Away_FIP = if_else(Home == 1, lag(FIP), FIP),
         Home_SOPerc = if_else(Home == 0, lead(SOPerc), SOPerc),
         Away_SOPerc = if_else(Home == 1, lag(SOPerc), SOPerc),
         Home_SOPercH = if_else(Home == 0, lead(SOPercH), SOPercH),
         Away_SOPercH = if_else(Home == 1, lag(SOPercH), SOPercH),
         Home_K9 = if_else(Home == 0, lead(K_9), K_9),
         Away_K9 = if_else(Home == 1, lag(K_9), K_9),
         Home_KBB = if_else(Home == 0, lead(K_BB), K_BB),
         Away_KBB = if_else(Home == 1, lag(K_BB), K_BB),
         Home_HR9 = if_else(Home == 0, lead(HR_9), HR_9),
         Away_HR9 = if_else(Home == 1, lag(HR_9), HR_9),
         Home_RunDiff = if_else(Home == 0, lead(Run_Diff), Run_Diff))%>%
  select(Date, Matchup, Home_Team, Home_BA, Home_OBP, Home_SLG, Home_OPS, Home_ISO, Home_SOPercH, Home_BAA, Home_OBPA, Home_SLGA, Home_OPSA, Home_StrPerc, Home_ERA, Home_WHIP, Home_FIP, Home_SOPerc, Home_KBB, Home_K9, Home_HR9, Away_Team, Away_BA, Away_OBP, Away_SLG, Away_OPS, Away_ISO, Away_SOPercH, Away_BAA, Away_OBPA, Away_SLGA, Away_OPSA, Away_StrPerc, Away_ERA, Away_WHIP, Away_FIP, Away_SOPerc, Away_KBB, Away_K9, Away_HR9, Home_Win, Home_RunDiff)%>%
  unique
```



```{r}
View(games.train2)
```


```{r}
games.train3 <- games.train2 %>%
  select(4:21, 23:41)

games.test3 <- games.test2 %>%
  select(4:21, 23:41)

mod.log.cumu <- glm(Home_Win ~.,
               data = games.train3,
               family = "binomial")

games.test.log <- games.test3 %>%
  add_predictions(mod.log.cumu,
                  type = "response") %>%
  mutate(class = if_else(pred > 0.5, 1,0)) #class instead of response

table(games.test.log$Home_Win, games.test.log$class)
(err.log.cumu <- with(games.test.log, mean(Home_Win != class)))
```

```{r}
vip(mod.log.cumu)
```

```{r}
res.vip4 <- vip(mod.log.cumu,num_features=55)
vip.dat4 <- res.vip4[["data"]]
vip.dat4[29:32,]
```

```{r}
summary(mod.log.cumu)
```



```{r}
mod.lm.cumu <- lm(Home_Win ~.,
               data = games.train3)
```

```{r}
games.test.lm <- games.test3 %>%
  add_predictions(mod.lm.cumu,
                  type = "response") %>%
  mutate(class = if_else(pred > 0.5, 1,0))
```

```{r}
table(games.test.lm$Home_Win, games.test.lm$class)
(err.lm.cumu <- with(games.test.lm, mean(Home_Win != class)))
```

```{r}
vip(mod.lm.cumu)
```

```{r}
res.vip5 <- vip(mod.lm.cumu,num_features=55)
vip.dat5 <- res.vip5[["data"]]
vip.dat5[29:32,]
```

```{r}
summary(mod.lm.cumu)
```


```{r}
games.y3 <- data.matrix(games.train2$Home_Win)


##New data set, gotta adjust the numbers selected
games.x3 <- games.train2 %>%
  select(4:21, 23:40)
games.x3[is.na(games.x3)] <- 0
games.x3 <- data.matrix(games.x3)
```

```{r}
ridge.cv.cumu <- cv.glmnet(games.x3, games.y3,
                      family="binomial",
                      type.measure="class", 
                      alpha=0)
plot(ridge.cv.cumu)
```


```{r}
log(ridge.cv.cumu$lambda.1se)
log(ridge.cv.cumu$lambda.1se)
```

```{r}
lambda.opt.cumu <- ridge.cv.cumu$lambda.1se
idrcumu <- with(ridge.cv.cumu,which(ridge.cv.cumu$lambda==lambda.opt.cumu))
(err.ridge.cumu <- ridge.cv.cumu$cvm[idrcumu])
```

```{r}
preds.ridge.cumu <- predict(ridge.cv.cumu, newx=games.test.x3,
                 type="class")
##The error rate on the testing dataset
table(games.test.y3 ,preds.ridge.cumu)
```

```{r}
vip(ridge.cv.cumu)
```

```{r}
res.vip6 <- vip(ridge.cv.cumu,num_features=55)
vip.dat6 <- res.vip6[["data"]]
vip.dat6[29:32,]
```


```{r}
mod.ridge.opt.cumu <- glmnet(games.x3, games.y3,
                        family="binomial",
                        type.measure="class",
                        alpha=0,
                        lambda=lambda.opt.cumu)

##New data set, gotta adjust the numbers selected
games.test.y3 <- data.matrix(games.test2$Home_Win)
games.test.x3 <- games.test2 %>%
  select(4:21, 23:40)
games.test.x3[is.na(games.test.x3)] <- 0
games.test.x3 <- data.matrix(games.test.x3)
```

```{r}
preds.rcumu <- predict(mod.ridge.opt.cumu, newx=games.test.x3,
                 type="class")
##The error rate on the testing dataset
table(games.test.y3 ,preds.rcumu)
```

```{r}
(err.ridge.opt.cumu <- mean((games.test.y3 != preds.rcumu)))
```

```{r}
vip(mod.ridge.opt.cumu)
```

```{r}
res.vip7 <- vip(mod.ridge.opt.cumu,num_features=55)
vip.dat7 <- res.vip7[["data"]]
vip.dat7[29:32,]
```



```{r}
games.y3 <- data.matrix(games.train2$Home_Win)


##New data set, gotta adjust the numbers selected
games.x.lasso <- games.train2 %>%
  select(4:21, 23:40)
games.x.lasso[is.na(games.x.lasso)] <- 0
games.x3 <- data.matrix(games.x.lasso)
```

```{r}
lasso.cv.cumu <- cv.glmnet(games.x3,
                      games.y3,
                      family="binomial",
                      type.measure="class", 
                      alpha=1)

plot(lasso.cv.cumu)
```

```{r}
vip(lasso.cv.cumu)
```

```{r}
preds.lcumu <- predict(lasso.cv.cumu, newx=games.test.x3,
                 type="class")
##The error rate on the testing dataset
table(games.test.y3 ,preds.lcumu)
```

```{r}
log(lasso.cv.cumu$lambda.1se)
log(lasso.cv.cumu$lambda.1se)
```

```{r}
lambda.opt.l <- lasso.cv.cumu$lambda.1se
idlcumu <- with(lasso.cv.cumu,which(lasso.cv.cumu$lambda==lambda.opt.l))
(err.lasso.cumu <- lasso.cv.cumu$cvm[idlcumu])
```


```{r}
mod.lasso.opt.cumu <- glmnet(games.x3, games.y3,
                        family="binomial",
                        type.measure="class",
                        alpha=1,
                        lambda=lambda.opt.l)
```

```{r}
preds <- predict(mod.lasso.opt, newx=games.test.x,
                 type="class")
##The error rate on the testing dataset
table(games.test.y ,preds)
```

```{r}
(err.lasso.opt <- mean((games.test.y != preds)))
```


```{r}
vip(mod.lasso.opt)
```


```{r}
games.train4 <- games.train2%>%
  mutate(BA_diff = Home_BA - Away_BA,
         OBP_diff = Home_OBP - Away_OBP,
         SLG_diff = Home_SLG - Away_SLG,
         OPS_diff = Home_OPS - Away_OPS,
         ISO_diff = Home_ISO - Away_ISO,
         SOPercH_diff = Home_SOPercH - Away_SOPercH,
         BAA_diff = Home_BAA - Away_BAA,
         OBPA_diff = Home_OBPA - Away_OBPA,
         SLGA_diff = Home_SLGA - Away_SLGA,
         OPSA_diff = Home_OPSA - Away_OPSA,
         StrPerc_diff = Home_StrPerc - Away_StrPerc,
         ERA_diff = Home_ERA - Away_ERA,
         WHIP_diff = Home_WHIP - Away_WHIP,
         FIP_diff = Home_FIP - Away_FIP,
         SOPerc_diff = Home_SOPerc - Away_SOPerc,
         KBB_diff = Home_KBB - Away_KBB,
         K9_diff = Home_K9 - Away_K9,
         HR9_diff = Home_HR9 - Away_HR9)

games.train5 <- games.train4%>%
  select(Date, Matchup, Home_Team, Away_Team, BA_diff, OBP_diff, SLG_diff, OPS_diff, ISO_diff, SOPercH_diff, BAA_diff, OBPA_diff, OPSA_diff, StrPerc_diff, ERA_diff, WHIP_diff, FIP_diff, SOPerc_diff, KBB_diff, K9_diff, HR9_diff, Home_Win, Home_RunDiff)

games.train6 <- games.train4%>%
  mutate(BA_percdiff = BA_diff/Home_BA,
         OBP_percdiff = OBP_diff/Home_OBP,
         SLG_percdiff = SLG_diff/Home_SLG,
         OPS_percdiff = OPS_diff/Home_OPS,
         ISO_percdiff = ISO_diff/Home_ISO,
         SOPercH_percdiff = SOPercH_diff/Home_SOPercH,
         BAA_percdiff = BAA_diff/Home_BAA,
         OBPA_percdiff = OBPA_diff/Home_OBPA,
         SLGA_percdiff = SLGA_diff/Home_SLGA,
         OPSA_percdiff = OPSA_diff/Home_OPSA,
         StrPerc_percdiff = StrPerc_diff/Home_StrPerc,
         ERA_percdiff = ERA_diff/Home_ERA,
         WHIP_percdiff = WHIP_diff/Home_WHIP,
         FIP_percdiff = FIP_diff/Home_FIP,
         SOPerc_percdiff = SOPerc_diff/Home_SOPerc,
         KBB_percdiff = KBB_diff/Home_KBB,
         K9_percdiff = K9_diff/Home_K9,
         HR9_percdiff = HR9_diff/Home_HR9)%>%
  select(Date, Matchup, Home_Team, Away_Team, BA_percdiff, OBP_percdiff, SLG_percdiff, OPS_percdiff, ISO_percdiff, SOPercH_percdiff, BAA_percdiff, OBPA_percdiff, OPSA_percdiff, StrPerc_percdiff, ERA_percdiff, WHIP_percdiff, FIP_percdiff, SOPerc_percdiff, KBB_percdiff, K9_percdiff, HR9_percdiff, Home_Win, Home_RunDiff)
  
```


```{r}
games.test4 <- games.test2%>%
  mutate(BA_diff = Home_BA - Away_BA,
         OBP_diff = Home_OBP - Away_OBP,
         SLG_diff = Home_SLG - Away_SLG,
         OPS_diff = Home_OPS - Away_OPS,
         ISO_diff = Home_ISO - Away_ISO,
         SOPercH_diff = Home_SOPercH - Away_SOPercH,
         BAA_diff = Home_BAA - Away_BAA,
         OBPA_diff = Home_OBPA - Away_OBPA,
         SLGA_diff = Home_SLGA - Away_SLGA,
         OPSA_diff = Home_OPSA - Away_OPSA,
         StrPerc_diff = Home_StrPerc - Away_StrPerc,
         ERA_diff = Home_ERA - Away_ERA,
         WHIP_diff = Home_WHIP - Away_WHIP,
         FIP_diff = Home_FIP - Away_FIP,
         SOPerc_diff = Home_SOPerc - Away_SOPerc,
         KBB_diff = Home_KBB - Away_KBB,
         K9_diff = Home_K9 - Away_K9,
         HR9_diff = Home_HR9 - Away_HR9)

games.test5 <- games.test4%>%
  select(Date, Matchup, Home_Team, Away_Team, BA_diff, OBP_diff, SLG_diff, OPS_diff, ISO_diff, SOPercH_diff, BAA_diff, OBPA_diff, OPSA_diff, StrPerc_diff, ERA_diff, WHIP_diff, FIP_diff, SOPerc_diff, KBB_diff, K9_diff, HR9_diff, Home_Win, Home_RunDiff)

games.test6 <- games.test4%>%
  mutate(BA_percdiff = BA_diff/Home_BA,
         OBP_percdiff = OBP_diff/Home_OBP,
         SLG_percdiff = SLG_diff/Home_SLG,
         OPS_percdiff = OPS_diff/Home_OPS,
         ISO_percdiff = ISO_diff/Home_ISO,
         SOPercH_percdiff = SOPercH_diff/Home_SOPercH,
         BAA_percdiff = BAA_diff/Home_BAA,
         OBPA_percdiff = OBPA_diff/Home_OBPA,
         SLGA_percdiff = SLGA_diff/Home_SLGA,
         OPSA_percdiff = OPSA_diff/Home_OPSA,
         StrPerc_percdiff = StrPerc_diff/Home_StrPerc,
         ERA_percdiff = ERA_diff/Home_ERA,
         WHIP_percdiff = WHIP_diff/Home_WHIP,
         FIP_percdiff = FIP_diff/Home_FIP,
         SOPerc_percdiff = SOPerc_diff/Home_SOPerc,
         KBB_percdiff = KBB_diff/Home_KBB,
         K9_percdiff = K9_diff/Home_K9,
         HR9_percdiff = HR9_diff/Home_HR9)%>%
  select(Date, Matchup, Home_Team, Away_Team, BA_percdiff, OBP_percdiff, SLG_percdiff, OPS_percdiff, ISO_percdiff, SOPercH_percdiff, BAA_percdiff, OBPA_percdiff, OPSA_percdiff, StrPerc_percdiff, ERA_percdiff, WHIP_percdiff, FIP_percdiff, SOPerc_percdiff, KBB_percdiff, K9_percdiff, HR9_percdiff, Home_Win, Home_RunDiff)
  
```



































